---
title: "Homework2"
author: "Savita Sastry"
date: "1/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("tidyverse")
```

```{r}
?msleep
#Proceedings of the National Academy of Sciences, 104 (3):1051-1056, 2007.
```

```{r}
#data frame called sleep, which calls the dataset msleep from ggplot2
sleep <- data.frame(ggplot2::msleep)
sleep
```
```{r}
#list of the column names from sleep data frame
colnames(sleep)
```

```{r}
glimpse(sleep)
```

```{r}
summary(sleep)
```

Print out the first column of the data frame and then print out the first row.
```{r}

sleep[,1]
sleep[1,]
```
We are interested in two groups; small and large mammals. Let’s define small as less than or equal to 1kg body weight and large as greater than or equal to 200kg body weight. Make two new dataframes (large and small) based on these parameters.
```{r}
  
small <- sleep[sleep$bodywt <= 1,]
large <- sleep[sleep$bodywt >= 200,]
small
large
```
What is the mean weight for both the large and small mammals?

```{r}
mean(small$bodywt) #mean weight in kg for mammals classified as small
mean(large$bodywt) #mean weight in kg for mammals classified as large
```
Let’s try to figure out if large mammals sleep, on average, longer than small mammals. What is the average sleep duration for large mammals as we have defined them? What is the average sleep duration for small mammals as we have defined them?
```{r}
mean(small$sleep_total) #mean sleep duration in hoursfor mammals classified as small
mean(large$sleep_total) #mean sleep duration in hoursfor mammals classified as large

#Conclisions from the above show that small mammals sleep longer than large mammals on average
```

Which animals are the sleepiest? Which sleep least 18 hours per day?
```{r}
sleepy_mammal <- sleep[sleep$sleep_total >= 18,]
sleepy_mammal

#sleepy_mammal dataframe displays all mammals that sleep over 18 hrs total or more per day.
```
















