---
title: "Homework 4"
author: "Savita Sastry"
date: "2/4/2020"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

```{r}
getwd()
```

```{r}
fisheries <- readr::read_csv("C:/Users/Savita/Documents/GitHub/BIS15W2020_ssastry/lab4/data/FAO_1950to2012_111914.csv")

fisheries
```

2. What are the names of the columns? Do you see any potential problems with the column names?
```{r}
colnames(fisheries)

#Problems with the current column names include it having all the years as columns as well as other stratifications that are not dates. Looks like untidy data
```

3. What is the structure of the data? Show the classes of each variable.

```{r}
str(fisheries)
```

```{r}
sapply(fisheries, class)
```

4. Think about the classes. Will any present problems? Make any changes you think are necessary below.

```{r}
#change values in the dates columns from character to numeric
fisheries[9:71] <- lapply(fisheries[9:71], as.numeric)
sapply(fisheries, class)

```

5. How many countries are represented in the data? Provide a count.

```{r}
#There are 204 countries represented in this data set
length(unique(fisheries$Country))

```

6. What are the names of the countries?

```{r}
unique(fisheries$Country)
```


7. Use `rename()` to rename the columns and make them easier to use. The new column names should be:  
+ country
+ commname
+ ASFIS_sciname
+ ASFIS_spcode
+ ISSCAAP_spgroup
+ ISSCAAP_spgroupname
+ FAO_area
+ unit

```{r}
fisheries <- fisheries %>% 
  dplyr::rename(
    country = "Country",
    commname = "Common name", 
    ASFIS_spcode = "ASFIS species#",
    ASFIS_sciname = "ASFIS species name",  
    ISSCAAP_spgroup = "ISSCAAP group#",
    ISSCAAP_spgroupname = "ISSCAAP taxonomic group", 
    FAO_area = "FAO major fishing area", 
    unit = "Measure")
      
fisheries
```

8. Are these data tidy? Why or why not, and, how do you know? Use the appropriate pivot function to tidy the data. Remove the NA's. Put the tidy data frame into a new object `fisheries_tidy`.  

Requirements for tidyness: 
(1) each variable has its own column  
(2) each observation has its own row  
(3) each value has its own cell  

The second condition is not met

```{r}

fisheries_tidy <- fisheries %>% 
  pivot_longer(`1950`:`2012`,
               names_to = "year", 
               values_to = "catch",
               values_drop_na = TRUE
               )

fisheries_tidy

```

9. Refocus the data only to include country, ISSCAAP_spgroupname, ASFIS_spcode, ASFIS_sciname, year, and catch.

```{r}
fisheries_tidy2 <- fisheries_tidy %>%
  select(country, ISSCAAP_spgroupname, ASFIS_spcode, ASFIS_sciname, year, catch)

fisheries_tidy2
```

10. Re-check the classes of `fisheries_tidy2`. Notice that "catch" is shown as a character! This is a problem because we will want to treat it as a numeric. How will you deal with this? 

```{r}
sapply(fisheries_tidy2, class)
```

```{r}
fisheries_tidy2$catch <- as.numeric(fisheries_tidy2$catch)
sapply(fisheries_tidy2, class)
```

11. Based on the ASFIS_spcode, how many distinct taxa were caught as part of these data?

```{r}
length(unique(fisheries_tidy2$ASFIS_spcode))

#1530 distinct taxa
```
 

12. Which country had the largest overall catch in the year 2000?

```{r}
catch_by_country <- fisheries_tidy %>%
  select(country, catch, year) %>%
  filter(year == 2000) %>%
  group_by(country) %>%
  summarize(catch_total = sum(catch, na.rm = T)) %>%
  arrange(desc(catch_total))

catch_by_country

#Peru had the largest number of catches
```
 
13. Which country caught the most sardines (_Sardina_) between the years 1990-2000?

```{r}
fisheries_tidy2 %>%
  select(country, year, ASFIS_sciname, catch) %>%
  filter(str_detect(ASFIS_sciname, "Sardina")) %>%
  filter(year >= "1990" & year <= "2000") %>%
  group_by(country) %>%
  summarize(catch = sum(catch)) %>%
  arrange(desc(catch))
 
#Morocco caught the most sardines
  
```

14. Which five countries caught the most cephalopods between 2008-2012?

```{r}
#cephalopods include Squids, cuttlefishes, octopuses

fisheries_tidy2 %>%
  select(country, catch, year, ISSCAAP_spgroupname) %>%
  filter(year >= "2008" & year <= "2012") %>%
  filter(str_detect(ISSCAAP_spgroupname, "Squids, cuttlefishes, octopuses")) %>%
  group_by(country) %>%
  summarize(catch = sum(catch)) %>%
  arrange(desc(catch))

#Top 5 countries: China, Peru, Korea, Japan, Chile
```

15. Given the top five countries from question 12 above, which species was the highest catch total? The lowest?

```{r}

fisheries_tidy2 %>%
  filter(str_detect(ISSCAAP_spgroupname, "Squids, cuttlefishes, octopuses")) %>% 
  filter(year >= "2008" & year <= "2012") %>%
  group_by(ASFIS_sciname) %>%
  summarize(catch_total = sum(catch, na.rm = T)) %>%
  arrange((catch_total))
      
  
```


16. In some cases, the taxonomy of the fish being caught is unclear. Make a new data frame called `coastal_fish` that shows the taxonomic composition of "Miscellaneous coastal fishes" within the ISSCAAP_spgroupname column.

```{r}
coastal_fish <- fisheries_tidy2 %>%
  filter(str_detect(ISSCAAP_spgroupname, "Miscellaneous coastal fishes"))
 
coastal_fish
```



17. Use the data to do at least one exploratory analysis of your choice. What can you learn?

```{r}
#investigate which species are the most popular in catch number worldwide.
#Identify the species which is caught the most and look at its distribution per country


popular_fish <- fisheries_tidy2 %>%
 group_by(ASFIS_sciname) %>%
  summarize(catch_total = sum(catch, na.rm = T)) %>%
  filter(catch_total >= 50000000) %>%
  arrange(desc(catch_total))

popular_fish

```

```{r}

#The osteichthyes, or bony fish, represent the largest taxonomic class of vertebrates in the modern world.

most_caught <- fisheries_tidy2 %>%
  filter(str_detect(ASFIS_sciname, "Osteichthyes")) %>%
  group_by(country) %>%
  summarize(catch = sum(catch)) %>%
  arrange(desc(catch))
 
most_caught
```
```{r}
#Visualize above data

#install.packages("rworldmap")
#library(rworldmap)

```

```{r}
# matched <- joinCountryData2Map(most_caught, joinCode="NAME", nameJoinColumn="country")
# 
# mapCountryData(matched, nameColumnToPlot="catch", mapTitle="Catch Country Sample", catMethod = "pretty", colourPalette = "heat")

```

