---
title: "Homework 6"
author: "Savita Sastry"
date: "2/18/2020"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

```{r}
#install.packages("gapminder")
library("gapminder")
```

```{r}
library(ggplot2)
```



**1. Use the function(s) of your choice to get an idea of the overall structure of the data frame, including its dimensions, column names, variable classes, etc. As part of this, determine how NAs are treated in the data.**  

```{r}
gapminder <- gapminder
```


```{r}
dim(gapminder)
```

```{r}
colnames(gapminder)
```

```{r}
str(gapminder)
```

```{r}
summary(gapminder)
```

```{r}
gapminder %>%
  summarize(na_count = sum(is.na(gapminder)))
```


**2. Among the interesting variables in gapminder is life expectancy. How has global life expectancy changed between 1952 and 2007?**

```{r}
gapminder %>%
  group_by(year) %>%
  summarize(avg_life_exp = mean(lifeExp)) %>%
  ggplot(aes(x = year, y = avg_life_exp)) +
  geom_line()+
  geom_point() +
  xlab("Year")  +
  ylab("Average Life Expectancy") +
  labs(title = "Life Expectancy per Year")

```


**3. How do the distributions of life expectancy compare for the years 1952 and 2007? _Challenge: Can you put both distributions on a single plot?_**
##This works absolutely fine! But note that you can skip a few of these coding steps and just filter the data by each of the two years and plot it against against life expectancy.
```{r}
life_1952 <- gapminder %>%
  select(lifeExp, year)%>%
  filter(year == 1952)

life_2007 <- gapminder %>%
  group_by(year) %>%
  select(lifeExp, year)%>%
  filter(year == 2007)
 
```

```{r}
life_2007

```

```{r}
life_1952
```

```{r}

life_exp_comparison <- ggplot(data = life_2007, aes(x= lifeExp))+
  geom_density(fill="steelblue", alpha  =0.6, color = "black") +
  # Change the fill colour to differentiate it
  geom_density(data= life_1952, fill="pink", alpha  =0.6, color = "black") +
  labs(title = "Distribution of Life Expectancy in 1957 and 2007")+
  labs(y="Density")+
  labs(x="Life Expectancy")

life_exp_comparison

#Results show that distribution of life expectancy between 1952 and 2007 has shifted considerably to the right (higher life expectancy)

```


**4. Your answer above doesn't tell the whole story since life expectancy varies by region. Make a summary that shows the min, mean, and max life expectancy by continent for all years represented in the data.**

```{r}
continent_lifeExp <- gapminder %>% 
  group_by(continent, year) %>% 
  summarize(min_lifeExp =min(lifeExp),
            max_lifeExp =max(lifeExp),
            mean_lifeExp =mean(lifeExp))

continent_lifeExp
```


**5. How has life expectancy changed between 1952-2007 for each continent? Try using `geom_line()` for this, including all continents on the same  plot.**

##I just love these point lines
```{r}
continent_lifeExp %>%
  ggplot(aes(x = year, y = mean_lifeExp, col = continent)) +
  geom_line()+
  geom_point() +
  xlab("Year")  +
  ylab("Average Life Expectancy") +
  labs(title = "Life Expectancy per Continent Through Time")
```

**6. We are interested in the relationship between per capita GDP and life expectancy; i.e. does having more money help you live longer?**

```{r}
gapminder %>%
  ggplot(aes(x = lifeExp, y = gdpPercap)) +
  geom_point() +
  geom_smooth(method=lm, se=F) + 
  xlab("Life Expectancy")  +
  ylab("per capita GDP") +
  labs(title = "Life Expectancy per capita GDP") 

#There seems to be some sort of positive correlation between life expectancy and per capita GDP
  
```


**7. There is extreme disparity in per capita GDP. Rescale the x axis to make this easier to interpret. How would you characterize the relationship?**
##This trims the data down, but to scale it we might want to use something like scale_x_log10() and geom_smoth() to make the relationship really pop out.
```{r}
gapminder %>%
  filter(gdpPercap <= 50000) %>%
  ggplot(aes(x = lifeExp, y = gdpPercap)) +
  geom_point() +
  geom_smooth(method=lm) + 
  xlab("Life Expectancy")  +
  ylab("per capita GDP") +
  labs(title = "Life Expectancy per capita GDP") 


```

**8. Which countries have had the largest population growth since 1952?**

```{r}
#population growth can be characterized as  the difference in population average size in 2007 vs average size in 1952

population_growth <- gapminder %>% 
  group_by(country, year) %>% 
  summarize(mean_pop =mean(pop)) %>%
  filter(year == 1952 | year == 2007) %>%
  pivot_wider(names_from = year, values_from = mean_pop) %>%
  dplyr::rename(
    Year_1952 = '1952',
    Year_2007 = '2007'
  ) %>%
  mutate(pop_grow = Year_2007 - Year_1952) %>%
  arrange(desc(pop_grow))

population_growth
```


**9. Use your results from the question above to plot population growth for the top five countries since 1952.**

```{r}
head(population_growth, 5)
```


```{r}
head(population_growth, 5) %>% 
  arrange(desc(pop_grow)) %>%
  ggplot(aes(x = country, y = pop_grow, fill = country)) +
  geom_bar(position = "dodge", color = "black", stat = "Identity") +
  xlab("Country")  +
  ylab("Population Growth") +
  labs(title = "Population Growth of Top 5 Countries Since 1952")
  
```


**10. How does per capita GDP growth compare between these same five countries?**


```{r}
gdp_country <- gapminder %>%
  group_by(country, year) %>% 
  summarize(mean_gdp = mean(gdpPercap)) %>%
  filter(year == 1952 | year == 2007) %>%
  pivot_wider(names_from = year, values_from = mean_gdp) %>%
  dplyr::rename(
    gdp_1952 = '1952',
    gdp_2007 = '2007') %>%
  mutate(gdp_change = gdp_2007 - gdp_1952) %>%
  arrange(desc(gdp_change))

gdp_country
```


```{r}
gdp_country %>% 
  filter(country == "Brazil" | country == "China" | country == "India" | country == "Indonesia" | country == "United States") %>%
  ggplot(aes(x = country, y = gdp_change, fill = country)) +
  geom_bar(position = "dodge", color = "black", stat = "Identity") +
  xlab("Country")  +
  ylab("Per Capita GDP") +
  labs(title = "GDP per Capita Growth of Top 5 Countries Since 1952")

#Population growth and per capita gdp do not reflect the same way in these top 5 countries
```

##Bars work well enough to show the difference between countries, but maybe if we're interested in plotting growth over time we might consider a line graph, yeah?